HTW Modeling

### 2024-03-23 Model/htw_model.qmd
added e2 & e3 brms error model tables. Reorganized plots

```{r}
pacman::p_load(dplyr,purrr,tidyr,ggplot2, data.table, here, patchwork, conflicted, 
               stringr,future,furrr, knitr, reactable, flextable,ggstance, htmltools,ggdist,ggh4x)
#conflict_prefer_all("dplyr", quiet = TRUE)
walk(c("flextable","dplyr"), conflict_prefer_all, quiet = TRUE)
#options(brms.backend="cmdstanr",mc.cores=4)
options(digits=3, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("Display_Functions","fun_alm","fun_indv_fit","fun_model", "prep_model_data","org_functions"), ~source(here::here(paste0("Functions/", .x, ".R"))))

@@

```{r}
#| label: tbl-htw-modelError
#| tbl-cap: "Mean model errors predicting empirical data from the testing and training stage, aggregated over all participants and velocity bands. Note that Fit Method refers to the subset of the data that the model was trained on"
post_tabs <- abc_tables(post_dat,post_dat_l)
train_tab <- abc_train_tables(pd_train,pd_train_l)

rbind(post_tabs$agg_pred_full |> mutate(stage="Test"), train_tab$agg_pred_full |> mutate(stage="Train")) |> 
  mutate(Fit_Method=rename_fm(Fit_Method)) |>
  flextable::tabulator(rows=c("stage","Fit_Method","Model"), columns=c("condit"),
                       `ME` = as_paragraph(mean_error)) |> as_flextable()
 # post_dat  |> group_by(condit,Model,Fit_Method,x) |> 
 #    mutate(e2=abs(dist-pred_dist)) |> 
 #    summarise(dist=mean(dist), pred=mean(pred_dist), mean_error=mean(e2)) |>
 #    group_by(condit,Model,Fit_Method) |> 
 #    summarise(mean_error=mean(mean_error)) |> 
 #    round_tibble(1) |> 
 #  mutate(Fit_Method=rename_fm(Fit_Method)) |>
 #  flextable::tabulator(rows=c("Fit_Method","Model"), columns=c("condit"),
 #                       `ME` = as_paragraph(mean_error)) |> as_flextable()

@@

library(brms)
options(brms.backend="cmdstanr",mc.cores=4)

@@

bct <- as.data.frame(bayestestR::describe_posterior(e1_ee_brm_ae, centrality = "Mean"))[, c(1,2,4,5,6)]
colnames(bct) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")

bct <- bct |> mutate(across(where(is.numeric), \(x) round(x, 2))) |>
  tibble::remove_rownames() |>
  mutate(Term = stringr::str_remove(Term, "b_")) |> kable(booktabs = TRUE)

@@

p1 + p2+ p3

@@

```{r}
#| label: fig-cm-dev-pat
#| fig-cap: "Empirical data and Model predictions for absolute deviation from target (accuracy). Observed data is shown on top, and predictions from each combination of model and fitting method are shown below. "
#| fig-height: 11
#| fig-width: 11

@@

# pmod1 <- post_dat_l |> filter(!(Resp=="Observed")) |> 
#   left_join(testAvgE1, by=join_by(id,condit,x==bandInt)) |>
#   group_by(id,condit, Fit_Method,Resp,x,bandType) |> 
#   mutate(x=as.factor(x)) |>
#  summarize(dist=median(dist)) |> 
#   mutate(oc = ifelse(bandType == "Trained", "black", NA)) |> 
#   ggplot(aes(x=condit,y=dist, fill=x,col=oc,size=oc)) + 
#   stat_bar +
#   ggh4x::facet_nested_wrap(~Resp+rename_fm(Fit_Method), axes = "all",ncol=3) + 
#    scale_color_manual(values = c("black" = "black"), guide = FALSE) +
#    scale_size_manual(values = c("black" = .3), guide = FALSE) + 
#   theme(legend.position = "none",plot.title = element_text(hjust=.55) ) + 
#   labs(title="Model Predictions", y="Deviation from target band", x="Condition")

@@

```{r}
#| fig-height: 11
#| fig-width: 11

@@

rbind(post_tabs2$agg_pred_full |> mutate(stage="Test"), train_tab2$agg_pred_full |> mutate(stage="Train")) |> 
  mutate(Fit_Method=rename_fm(Fit_Method)) |>
  flextable::tabulator(rows=c("stage","Fit_Method","Model"), columns=c("condit"),
                       `ME` = as_paragraph(mean_error)) |> as_flextable()

@@

bct <- as.data.frame(bayestestR::describe_posterior(e2_ee_brm_ae, centrality = "Mean"))[, c(1,2,4,5,6)]
colnames(bct) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")

bct <- bct |> mutate(across(where(is.numeric), \(x) round(x, 2))) |>
  tibble::remove_rownames() |>
  mutate(Term = stringr::str_remove(Term, "b_")) |> kable(booktabs = TRUE)

p1 <- plot(conditional_effects(e2_ee_brm_ae, effects="condit"),points=FALSE, plot=FALSE)$condit + 
  ggplot2::xlab("Condition") +ylab("Model Error")
p2 <- plot(conditional_effects(e2_ee_brm_ae, effects="Model"),points=FALSE, plot=FALSE)$Model + 
  labs(x="Model",y=NULL)
p3 <- plot(conditional_effects(e2_ee_brm_ae, effects="Model:condit"),points=FALSE, plot=FALSE)$`Model:condit` + 
  scale_color_manual(values=wes_palette("Darjeeling1")) +
  labs(x="Model",y=NULL,fill=NULL,col=NULL) + theme(legend.position="right")

 p1 + p2+ p3

pemp2 <- e2 |> filter(expMode2=="Test") |> mutate(Resp="Observed") |> 
  ggplot(aes(x=condit,y=vx, fill=vb, col=ifelse(bandType=="Trained","black",NA),size=ifelse(bandType=="Trained","black",NA))) + 
  stat_bar+ 
  scale_color_manual(values = c("black" = "black"), guide = FALSE) +
  scale_size_manual(values = c("black" = .5), guide = FALSE) + 
  theme(legend.position="right", axis.title.x = element_blank(), plot.title = element_text(hjust=.50)) +
  labs(title="Empirical Data - Experiment 1",y="vx", x="Condition",fill="Band")

pmod2 <- e2_model$post_dat_l |> filter(!(Resp=="Observed")) |> 
  group_by(id,condit, Fit_Method,Resp,bandType,x,vb) |> 
 summarize(vx=median(val)) |> 
 ggplot( aes(x=condit,y=vx, fill=vb,col=ifelse(bandType=="Trained","black",NA),size=ifelse(bandType=="Trained","black",NA))) + 
  stat_bar + 
    facet_wrap(~Resp+rename_fm(Fit_Method), strip.position = "top", scales = "free_x") +
        scale_color_manual(values = c("black" = "black"), guide = "none") +
  scale_size_manual(values = c("black" = .5), guide = "none") +
    theme(panel.spacing = unit(0, "lines"), 
         strip.background = element_blank(),
         strip.placement = "outside",
         legend.position = "none",plot.title = element_text(hjust=.50),
         axis.title.x = element_blank(),
         plot.margin = unit(c(20,0,0,0), "pt")) + 
         labs(title="Model Predictions - Experiment 1 Data", y="vx")

(pemp2)  / pmod2 + 
  plot_layout(design = layout, heights = unit(c(4, 9), c('cm','null')))
```

e3_model <- load_e3()
post_tabs3 <- abc_tables(e2_model$post_dat,e3_model$post_dat_l)
train_tab3 <- abc_train_tables(e2_model$pd_train,e3_model$pd_train_l)

@@

rbind(post_tabs3$agg_pred_full |> mutate(stage="Test"), train_tab3$agg_pred_full |> mutate(stage="Train")) |> 
  mutate(Fit_Method=rename_fm(Fit_Method)) |>
  flextable::tabulator(rows=c("stage","Fit_Method","Model"), columns=c("condit"),
                       `ME` = as_paragraph(mean_error)) |> as_flextable()

@@

bct <- as.data.frame(bayestestR::describe_posterior(e3_ee_brm_ae, centrality = "Mean"))[, c(1,2,4,5,6)]
colnames(bct) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")

bct <- bct |> mutate(across(where(is.numeric), \(x) round(x, 2))) |>
  tibble::remove_rownames() |>
  mutate(Term = stringr::str_remove(Term, "b_")) |> kable(booktabs = TRUE)

@@

plot(conditional_effects(e3_ee_brm_ae, 
                         effects = "Model:condit", 
                         conditions=make_conditions(e3_ee_brm_ae,vars=c("bandOrder"))),
     points=FALSE,plot=TRUE)

@@

 p1 + p2+ p3

@@

>>>>

```{r}
pacman::p_load(dplyr,purrr,tidyr,ggplot2, data.table, here, patchwork, conflicted, 
               stringr,future,furrr, knitr, reactable, flextable,ggstance, htmltools,ggdist,ggh4x,brms,tidybayes,bayestestR, gt)
#conflict_prefer_all("dplyr", quiet = TRUE)
walk(c("flextable","dplyr"), conflict_prefer_all, quiet = TRUE)
#options(brms.backend="cmdstanr",mc.cores=4)
options(digits=3, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("Display_Functions","fun_alm","fun_indv_fit","fun_model", "prep_model_data","org_functions"), ~source(here::here(paste0("Functions/", .x, ".R"))))

@@

```{r}
#| label: tbl-htw-modelError-e1
#| tbl-cap: "Mean model errors predicting empirical data from the testing and training stage, aggregated over all participants and velocity bands. Note that Fit Method refers to the subset of the data that the model was trained on"
post_tabs <- abc_tables(post_dat,post_dat_l)
train_tab <- abc_train_tables(pd_train,pd_train_l)

e1_tab <- rbind(post_tabs$agg_pred_full |> mutate("Task Stage"="Test"), train_tab$agg_pred_full |> mutate("Task Stage"="Train")) |> 
  mutate(Fit_Method=rename_fm(Fit_Method)) 
  # |>
  # flextable::tabulator(rows=c("stage","Fit_Method","Model"), columns=c("condit"),
  #                      `ME` = as_paragraph(mean_error)) |> as_flextable()

e1_tab %>%
  group_by(`Task Stage`, Fit_Method, Model, condit) %>%
  summarize(ME = mean(mean_error), .groups = "drop") %>%
  pivot_wider(
    names_from = c(Model, condit),
    values_from = ME,
    names_sep = "_"  # Add this line to specify the separator for column names
  ) %>%
  rename("Fit Method" = Fit_Method) %>%
  gt() %>%
  tab_header(
    title = "Mean Error by Fit Method, Task Stage, and Condition",
    subtitle = "Comparison across Models"
  ) %>%
  cols_move_to_start(columns = c(`Task Stage`)) %>%
  cols_label(
    `Task Stage` = "Task Stage"
  ) %>%
  fmt_number(
    columns = starts_with("ALM") | starts_with("EXAM"),
    decimals = 2
  ) %>%
  tab_spanner_delim(delim = "_") %>%
  tab_style(
    style = cell_fill(color = "white"),
     locations = cells_body(columns = everything(), rows = everything())
  ) %>%
  tab_style(
    style = cell_borders(sides = "top", color = "black", weight = px(1)),
    locations = cells_column_labels()
  ) %>%
  tab_options(
    column_labels.font.size = 10,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    table.font.size = 10, 
    quarto.disable_processing = TRUE
  ) %>%
  tab_footnote(
    footnote = "Note: Mean errors are calculated as the absolute difference between the predicted and actual values.",
    locations = cells_title(groups = "subtitle")
  ) %>%
  sub_missing(missing_text = "")

#  post_dat  |> group_by(condit,Model,Fit_Method,x) |> 
#     mutate(e2=abs(dist-pred_dist)) |> 
#     summarise(dist=mean(dist), pred=mean(pred_dist), mean_error=mean(e2)) |>
#     group_by(condit,Model,Fit_Method) |> 
#     summarise(mean_error=mean(mean_error)) |> 
#     round_tibble(1) |> 
#   mutate(Fit_Method=rename_fm(Fit_Method)) |>
#   flextable::tabulator(rows=c("Fit_Method","Model"), columns=c("condit"),
#                        `ME` = as_paragraph(mean_error)) |> as_flextable()

@@

# library(brms)
# options(brms.backend="cmdstanr",mc.cores=4)

@@

bct_e1 <- as.data.frame(bayestestR::describe_posterior(e1_ee_brm_ae, centrality = "Mean")) %>%
  select(1,2,4,5,6) %>%
  setNames(c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")) %>%
  mutate(across(where(is.numeric), \(x) round(x, 2))) %>%
  tibble::remove_rownames() %>%
  mutate(Term = stringr::str_remove(Term, "b_")) %>%
  kable(booktabs = TRUE)

@@

(p1 + p2+ p3) + plot_annotation(tag_levels = c('A'), tag_suffix=".")

@@

```{r}
#| label: fig-cm-dev-pat
#| fig-cap: "Empirical data and Model predictions for absolute deviation from target (accuracy). Observed data is shown on top, and predictions from each combination of model and fitting method are shown below. "
#| fig-height: 11
#| fig-width: 11
#| eval: false

@@

@@

```{r}
#| label: tbl-htw-modelError-e23

@@

e2_tab <- rbind(post_tabs2$agg_pred_full |> mutate("Task Stage"="Test"), train_tab2$agg_pred_full |> mutate("Task Stage"="Train")) |> 
  mutate(Fit_Method=rename_fm(Fit_Method)) 
  # |>
  # flextable::tabulator(rows=c("Task Stage","Fit_Method","Model"), columns=c("condit"),
  #                      `ME` = as_paragraph(mean_error)) |> as_flextable()

@@

bct_e2 <- as.data.frame(bayestestR::describe_posterior(e2_ee_brm_ae, centrality = "Mean")) %>%
  select(1,2,4,5,6) %>%
  setNames(c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")) %>%
  mutate(across(where(is.numeric), \(x) round(x, 2))) %>%
  tibble::remove_rownames() %>%
  mutate(Term = stringr::str_remove(Term, "b_")) %>%
  kable(booktabs = TRUE)

e3_model <- load_e3()
post_tabs3 <- abc_tables(e3_model$post_dat,e3_model$post_dat_l)
train_tab3 <- abc_train_tables(e3_model$pd_train,e3_model$pd_train_l)

@@

e3_tab <- rbind(post_tabs3$agg_pred_full |> 
  mutate("Task Stage"="Test"), train_tab3$agg_pred_full |> mutate("Task Stage"="Train")) |> 
  mutate(Fit_Method=rename_fm(Fit_Method)) 
  # |>
  # flextable::tabulator(rows=c("Task Stage","Fit_Method","Model"), columns=c("condit"),
  #                      `ME` = as_paragraph(mean_error)) |> as_flextable()

e23_tab <- rbind(e2_tab |> mutate(Exp="E2"), e3_tab |> mutate(Exp="E3")) 
# |> 
#   flextable::tabulator(rows=c("Task Stage","Fit_Method","Model"), columns=c("Exp", "condit"),
#                        `ME` = as_paragraph(mean_error)) |> flextable::as_flextable()

wide_data <- e23_tab %>%
  pivot_wider(
    names_from = c(Exp, Model, condit),
    values_from = mean_error,
    names_glue = "{Exp}_{Model}_{condit}"
  ) %>%
  arrange(Fit_Method, `Task Stage`)  # Sort for clarity

# Create a gt table with the wider data
gt_table <- wide_data %>%
  gt() %>%
  tab_header(
    title = "Mean Error by Fit Method, Task Stage, and Experiment",
    subtitle = "Comparison across Conditions and Models"
  ) %>%
  cols_move_to_start(columns = c(`Task Stage`)) %>% # Move Task Stage column to the start
  cols_label(
    `Task Stage` = "Task Stage"
  ) %>%
  fmt_number(
    columns = matches("E2|E3"),
    decimals = 1  # Adjust the number of decimal places as needed
  ) %>%
  tab_spanner_delim(delim = "_") %>%
  tab_style(
    style = cell_fill(color = "white"), # Ensure readability in grayscale
    locations = cells_body(columns = everything(), rows = everything())
  ) %>%
  tab_style(
    style = cell_borders(sides = "top", color = "black", weight = px(1)),
    locations = cells_body(columns = everything(), rows = everything())
  ) %>%
  tab_options(
    column_labels.font.size = 10,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    table.font.size = 10,
    quarto.disable_processing = TRUE
  ) %>%
  tab_footnote(
    footnote = "Note: Mean errors are calculated as the absolute difference between the predicted and actual values.",
    locations = cells_title(groups = "subtitle")
  )

gt_table

@@

bct_e3  <- as.data.frame(bayestestR::describe_posterior(e3_ee_brm_ae, centrality = "Mean")) %>%
  select(1,2,4,5,6) %>%
  setNames(c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")) %>%
  mutate(across(where(is.numeric), \(x) round(x, 2))) %>%
  tibble::remove_rownames() %>%
  mutate(Term = stringr::str_remove(Term, "b_")) %>%
  kable(booktabs = TRUE)

```{r}
#| label: fig-e2_e3_ae
#| fig-cap: "Conditional effects of Model and Condition on Model Error for Experiment 2 and 3 data."
#| fig-height: 11
#| fig-width: 11

p1 <- plot(conditional_effects(e2_ee_brm_ae, effects="condit"),points=FALSE, plot=FALSE)$condit + 
  ggplot2::xlab("Condition") +ylab("Model Error") + labs(title="E2. Model Error")
p2 <- plot(conditional_effects(e2_ee_brm_ae, effects="Model"),points=FALSE, plot=FALSE)$Model + 
  labs(x="Model",y=NULL)
p3 <- plot(conditional_effects(e2_ee_brm_ae, effects="Model:condit"),points=FALSE, plot=FALSE)$`Model:condit` + 
  scale_color_manual(values=wes_palette("Darjeeling1")) +
  labs(x="Model",y=NULL,fill=NULL,col=NULL) + theme(legend.position="right")

 p_e2 <- (p1 + p2+ p3)

@@

p_e3 <- plot(conditional_effects(e3_ee_brm_ae, 
                         effects = "Model:condit", 
                         conditions=make_conditions(e3_ee_brm_ae,vars=c("bandOrder"))),
     points=FALSE,plot=FALSE)$`Model:condit` + 
     labs(x="Model",y="Model Error", title="E3. Model Error") + 
     theme(legend.position="right") + 
     scale_color_manual(values=wes_palette("Darjeeling1"))

@@

 #p2 <- (p1 + p2+ p3)

@@

#modelsummary::modelsummary(list("e1"=e1_ee_brm_ae,"e2"=e2_ee_brm_ae,"e3"=e3_ee_brm_ae), gof_map = tribble())

 (p_e2 / p_e3) + plot_annotation(tag_levels = c('A'), tag_suffix=".")

```{r}
#| fig-height: 10
#| fig-width: 11
#| fig-cap: "Empirical data and Model predictions from Experiment 2 and 3 for the testing stage. Observed data is shown on the right. "

rbind(e2_model$post_dat_l |> filter( Fit_Method=="Test_Train") |> 
  group_by(id,condit, Fit_Method,Resp,bandType,x,vb) |> 
 summarize(vx=median(val)) |> mutate(Exp="E2",bandOrder="Reverse"), 
 e3_model$post_dat_l |> filter( Fit_Method=="Test_Train") |> 
  group_by(id,condit, Fit_Method,Resp,bandType,x,vb,bandOrder) |>
  summarize(vx=median(val)) |> mutate(Exp="E3")) |>
  ggplot( aes(x=condit,y=vx, fill=vb,col=ifelse(bandType=="Trained","black",NA),size=ifelse(bandType=="Trained","black",NA))) +
  stat_bar + 
    facet_nested_wrap(~Exp+bandOrder+Resp, strip.position = "top", scales = "free_x") +
    scale_color_manual(values = c("black" = "black"), guide = "none") +
  scale_size_manual(values = c("black" = .7), guide = "none") +
    theme(panel.spacing = unit(0, "lines"), 
        #  strip.background = element_blank(),
        #  strip.placement = "outside",
         legend.position = "none",plot.title = element_text(hjust=.50),
         axis.title.x = element_blank(),
         plot.margin = unit(c(20,0,0,0), "pt")) + 
         labs(title="Model Predictions Experiment 2 & 3", y="vx")
